---
title: "Custom ggplot themes"
author: "Galen Holt"
---

I often want to consistently theme my ggplots across projects. I've developed some custom themes, but they're usually ad-hoc, and don't work particularly well in packages, because the simple way to do it isn't a function.

```{r}
library(ggplot2)
```

For example, we might have a theme that's good for publication, as in [Saving and theming plots](tweaks_tricks.qmd), where we specify size, backgrounds, and text. We load new fonts first.

```{r}
# Load local functions
devtools::load_all()

showtext::showtext_auto()
pubfont <- 'Cambria'
loadfonts(fontvec = pubfont)
```

```{r}
pubtheme <- theme_bw(base_size = 10) + 
  theme(strip.background = element_blank(),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        text = element_text(family=pubfont))
```

Here, the theme is essentially hardcoded, and we can add it to a figure.

```{r}
ggplot(iris, aes(x = Sepal.Length, y = Petal.Width, color = Species)) + 
  geom_point() + 
  facet_wrap(~Species) + 
  pubtheme
```

### What's the problem?

First, we have to adjust the theme with `+` if we want to make changes.

Most importantly, probably, is that in packages that gets read in as an object, instead of a function (which yes, are objects, but packages like functions).

### Making it a function

The solution is to make it a function. This matches how the built-in themes work too, e.g. we add `theme_bw()`, not `theme_bw` , and that lets us modify it in the arguments. As a simple case, this lets us use the `base_size` to have a theme that we can give larger fonts to for talks vs papers, for example, whereas in the past I'd have two different names for that.

```{r}
theme_pub <- function(base_size = 10) {
  ggplot2::theme_bw(base_size = base_size) +
    theme(strip.background = element_blank(),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        text = element_text(family=pubfont))
}
```

This looks the same

```{r}
ggplot(iris, aes(x = Sepal.Length, y = Petal.Width, color = Species)) + 
  geom_point() + 
  facet_wrap(~Species) +
  theme_pub()
```

Now, maybe we also want to make the font an argument? Now, we can include `loadfonts` to load it if it's not already.

```{r}
theme_pubf <- function(base_size = 10, font) {
  if (!(font %in% sysfonts::font_families())) {
      loadfonts(font)
  }

  ggplot2::theme_bw(base_size = base_size) +
    theme(strip.background = element_blank(),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        text = element_text(family=font))
}
```

Now let's change the font

```{r}
ggplot(iris, aes(x = Sepal.Length, y = Petal.Width, color = Species)) + 
  geom_point() + 
  facet_wrap(~Species) +
  theme_pubf(base_size = 14, font = 'Calibri')
```

## Can we pass other arguments?

If the goal is to enforce a style, we might not want to *allow* passing other arguments to `theme`, but can we with `…`?

```{r}
theme_pub_dots <- function(base_size = 10, font, ...) {
  if (!(font %in% sysfonts::font_families())) {
      loadfonts(font)
  }

  ggplot2::theme_bw(base_size = base_size) +
    theme(strip.background = element_blank(),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        text = element_text(family=font), 
        ...)
}
```

```{r}
ggplot(iris, aes(x = Sepal.Length, y = Petal.Width, color = Species)) + 
  geom_point() + 
  facet_wrap(~Species) +
  theme_pub_dots(base_size = 8, font = 'Arial', legend.position = 'none')
```
