---
title: "Legend in missing facet"
format: html
---

I often want to use the existence of a 'missing' facet to put in a legend, and always have to re-figure out how, so I'm writing it down here.

```{r}
library(ggplot2)
library(patchwork)
```

Make a plot with a gap.

```{r}
irisplot <- iris |> 
  ggplot(aes(x = Sepal.Length, y = Petal.Width, color = Species)) +
  geom_point() +
  facet_wrap('Species', nrow = 2)

irisplot
```

We want the legend in that lower right hole. The easiest is to just use `theme(legend.position = c(x, y))` where x and y are in plot units.

```{r}
irisplot + theme(legend.position = c(0.6, 0.2))
```

Other options that can give more flexibility, particularly across plots, are to get the legend grob and treat it as another plot. That can't be plotted directly, but can with other plots. Because we can treat this as a plot, we can do some complex things with `ggpubr` and `patchwork`, including overlaying it into that spot or using it across figures or subsets of figures. Though here we just do something simple to demo.

`ggpubr` can use the grob directly

```{r}
ggpubr::ggarrange(irisplot + theme(legend.position = 'none'), irislegend)
```

patchwork needs to make it a ggplot

```{r}
(irisplot + theme(legend.position = 'none')) + ggpubr::as_ggplot(irislegend)
```

With patchwork we can use `inset_element` if we want to overlay and not treat as external plot

```{r}
(irisplot + theme(legend.position = 'none')) + 
  inset_element(ggpubr::as_ggplot(irislegend), left = 0.6, bottom = 0.2, right = 0.8, top = 0.4)
```
