---
title: "Math and greek in legends"
author: "Galen Holt"
output:
  html:
    extra_dependencies: ['amsmath']
editor: visual
---

```{r}
#| echo = FALSE

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

# The issue

I often need math in ggplot labels. Sometimes it's actual math, but often just things like greek letters, subscripts, fractions, hats, bars, etc. THere's ways to do most of that with `expression` and `paste` and `bquote`, but none of it is ever very intuitive for me. I'd like to be able to just send it latex and have it work.

Does `latex2exp` work?

## Using latex2exp

```{r}
#| message = FALSE,
#| warning = FALSE
library(ggplot2) 
library(latex2exp)
```

Let's just try to do some things with that. Make the usual `iris` plot.

```{r}
testplot <- ggplot(iris, aes(x = Sepal.Length, y = Petal.Width, color = Species)) +
  geom_point()
testplot
```

Now let's add some math. I'd typically use `labs` for the x,y, and color, so try that. Spacing not great, but it works.

Basically following [the manual](https://cran.r-project.org/web/packages/latex2exp/vignettes/using-latex2exp.html), it's pretty self-explanatory.

*The* `r` *says to use raw strings so don't have to escape slashes*.

```{r}
testplot <- testplot + 
  labs(x = TeX(r'(Words and greek $\Delta_1$)'),
       y = TeX(r'($\frac{1-\alpha}{\rho})'),
       color = TeX(r'($\left{ \int_0^\inf \exp{\eta x} dx \right})'))

testplot
```

Can I use `amsmath` in latex? Maybe, but not for linebreaks- this errors.

```{r}
#| error: false
#| eval: false
testplot <- testplot + 
  labs(x = TeX(r'(Words and greek $\Delta_1$)'),
       y = TeX(r'($\frac{1-\alpha}{\rho}$)'),
       color = TeX(r'($\begin{split}\left{ \int_0^\inf \\ \exp{\eta x} dx \right}\end{split}$)'))

testplot
```
