---
title: "Setting colors for groups"
author: "Galen Holt"
---

# The issue

I often want the same colors to map to the same levels across many plots within a project, even if those levels aren't included in a plot. E.g. if we plot the iris dataset with and without 'setosa', the colors change.

```{r}
library(ggplot2)
```

```{r}
ggplot(iris, aes(x = Sepal.Length, y = Petal.Width, color = Species)) +
  geom_point() +
  scale_color_brewer(palette = 'Dark2')
```

vs

```{r}
ggplot(dplyr::filter(iris, Species != 'setosa'), aes(x = Sepal.Length, y = Petal.Width, color = Species)) +
  geom_point() +
  scale_color_brewer(palette = 'Dark2')
```

Now, virginica and versicolor have changed colors, and that's confusing.

# Solution

We need to match colors. We can do this with `setNames`. It does require pre-allocating them and using a manual scale instead of something like `scale_color_brewer`, which feels hacky but in reality it makes sense- we *have* to know them all a-priori and assign them.

So, step 1 is to get the right number of colors from the desired palette and assign them. Most of the usual color palette packages, e.g. {colorspace}, {RColorBrewer}, {viridis}, {paletteer}, etc have ways of getting x colors from a given palette.

Colorbrewer is particularly weird syntax. I like paletteer and colorspace better.

```{r}
# Weird colorbrewer syntax
# iriscols <- scales::brewer_pal(palette = 'Dark2')(length(unique(iris$Species)))

# more intuitive colorspace syntax (which has palettes approximating brewer)
iriscols <- colorspace::qualitative_hcl(length(unique(iris$Species)), palette = 'Dark2')
```

## Assign colors

Now use `setNames` to match

```{r}
sppal <- setNames(iriscols, unique(iris$Species))
```

## Use manual scale and check

and re-do the above two plots to demonstrate that dropping levels keeps colors the same

```{r}
ggplot(iris, aes(x = Sepal.Length, y = Petal.Width, color = Species)) +
  geom_point() +
  scale_color_manual(values = sppal)
```

and drop setosa

```{r}
ggplot(dplyr::filter(iris, Species != 'setosa'), aes(x = Sepal.Length, y = Petal.Width, color = Species)) +
  geom_point() +
  scale_color_manual(values = sppal)
```

Now they match.

# Using in a package/project

For a project, we can do the `setNames` step, and then be consistent across plots to use `scale_color_manual()`. In a package that creates plots, we could do the same if we *know* that there will be a manual palette passed in. Otherwise we'd need to enforce it somehow or have a conditional that calls some other `scale_color…` type function if it's not there. There also might be a way to create our own `scale_color_custom()` function that handles it.

## Creating a scale\_ function

We should also probably provide a function to create the standard color-level matching. And maybe that's the solution to the above. If a palette isn't passed in, create one in-function.

First, can I make a silly wrapper that just takes the values? It's no different than scale_color_manual at this point

```{r}
scale_color_custom <- function(pal) {
  scale_color_manual(values = pal)
}
```

```{r}
ggplot(dplyr::filter(iris, Species != 'setosa'), aes(x = Sepal.Length, y = Petal.Width, color = Species)) +
  geom_point() +
  scale_color_custom(pal = sppal)
```

Now, can we catch times when that's not a named vector and if not pass something else? Do we want to? It starts getting infinite what we want to handle (e.g. palette names? from which packages?)

I suppose as long as it's a vector, use it but warn. I think otherwise just fail- it's too hard to catch everything. Something like

```{r}
scale_color_custom <- function(pal) {
  if (!is.character(pal)) {
    stop("pal needs to be a character vector, ideally named")
  }
  if (is.null(names(pal))) {
    warning("unnamed vector, colors may not be consistent between plots.")
  }
  
  scale_color_manual(values = pal)
}
```

## Palette generation

How about a palette generating function?

This will be easiest if I enforce a package, though I'm sure it can be made more flexible. For the sake of this quick demo, let's just use {paletteer}, since it has access to lots of options. And we at least are limited to discrete scales, since we're level-matching.

One thing I often want to do is set specific colors to specific levels (e.g. a reference). So make that possible.

```{r}
make_pal <- function(levels, palette, refvals = NULL, refcols = NULL) {
  if (is.factor(levels)) {levels <- as.character(levels)}
  nonrefs <- levels[!(levels %in% refvals)]
  cols <- paletteer::paletteer_d(palette, length(nonrefs))
  
  namedcols <- setNames(c(refcols, cols), c(refvals, nonrefs))
}
```

## Workflow

First, we set colors with `make_pal`

```{r}
ircols <- make_pal(unique(iris$Species), palette = 'calecopal::kelp2')
```

Then we plot with `scale_color_custom`

First with all three species

```{r}
ggplot(iris, aes(x = Sepal.Length, y = Petal.Width, color = Species)) +
  geom_point() +
  scale_color_custom(pal = ircols)
```

And removing setosa

```{r}
ggplot(dplyr::filter(iris, Species != 'setosa'), aes(x = Sepal.Length, y = Petal.Width, color = Species)) +
  geom_point() +
  scale_color_custom(pal = ircols)
```

### Reference level

Let's make setosa a reference level that stands out like purple.

```{r}
ircolsS <- make_pal(unique(iris$Species), palette = 'calecopal::kelp2', refvals = 'setosa', refcols = 'purple')
```

```{r}
ggplot(iris, aes(x = Sepal.Length, y = Petal.Width, color = Species)) +
  geom_point() +
  scale_color_custom(pal = ircolsS)
```

Note that now the colors of the others *have* shifted, because the reference-levelling happened at the start. If a reference is *always* a reference, that might be fine. But it might be the case that we only sometimes want to call attention to a particular level, and so we want to assign colors including it, so we can switch between an accentuated and unaccentuated palette. Let's put another argument in `make_pal` (which requires some reframing of how we assign things). We also might want to return both the unreferenced and the referenced palettes at the same time, rather than calling the function twice.

```{r}
make_pal <- function(levels, palette, refvals = NULL, refcols = NULL, includeRef = FALSE, returnUnref = FALSE) {
  
  if (returnUnref) {
    if (!includeRef) {
      stop("does not make sense to return a reffed and unreffed palette that don't match")
    }
    }
  
  if (is.factor(levels)) {levels <- as.character(levels)}
  
  if (!includeRef) {levels <- levels[!(levels %in% refvals)]}
  
  # nonrefs <- levels[!(levels %in% refvals)]
  cols <- paletteer::paletteer_d(palette, length(levels))
  
  if (returnUnref & includeRef) {unref <- setNames(cols, levels)}
  
  # delete the reference levels out of the vectors
  whichlevs <- which(!(levels %in% refvals))
  nonrefs <- levels[whichlevs]
  nonrefcols <- cols[whichlevs]
  
  namedcols <- setNames(c(refcols, nonrefcols), c(refvals, nonrefs))
  
  if (returnUnref & includeRef) {
    return(list(refcols = namedcols, unrefcols = unref))
    } else {
      return(namedcols)
  }

}
```

Now test that- setting includeref = TRUE should keep these the same as earlier plots except for setosa (i.e. the light blue that would go to setosa just drops out

```{r}
ircolsS_include <- make_pal(unique(iris$Species), palette = 'calecopal::kelp2', refvals = 'setosa', refcols = 'purple', includeRef = TRUE)
```

```{r}
ggplot(iris, aes(x = Sepal.Length, y = Petal.Width, color = Species)) +
  geom_point() +
  scale_color_custom(pal = ircolsS_include)
```

And if I want to get both palettes (reffed and unreffed) so I can accentuate sometimes,

```{r}
ircolsS_both <- make_pal(unique(iris$Species), palette = 'calecopal::kelp2', refvals = 'setosa', refcols = 'purple', includeRef = TRUE, returnUnref = TRUE)
```

With setosa as a reference- should match above

```{r}
ggplot(iris, aes(x = Sepal.Length, y = Petal.Width, color = Species)) +
  geom_point() +
  scale_color_custom(pal = ircolsS_both$refcols)
```

Without setosa as a reference, others retain same colors.

```{r}
ggplot(iris, aes(x = Sepal.Length, y = Petal.Width, color = Species)) +
  geom_point() +
  scale_color_custom(pal = ircolsS_both$unrefcols)
```
