---
title: "SVG vdiffr testing"
format: html
---

For some situations, I need to use golden tests for figures that *aren't* made by ggplot or base or otherwise have a `print()` method as required by `vdiffr::write_svg()`. Specifically, I encounter this issue with networks created by `DiagrammeR`, which are html. The solution is to write a custom `writer` function. Note that for this to work, we also need to install `DiagrammeRsvg` and `rsvg`; they are only suggests for `DiagrammeR` and so are not automatically installed.

```{r}
#| eval: false
svgnetwriter <- function(plot, file, title = "") {
  plot |>
    DiagrammeR::export_graph(file_name = paste0(file))
}
```

Then we can call `expect_doppleganger`

```{r}
#| eval: false
vdiffr::expect_doppelganger(title = "plot_name", fig = plot_object, 
                            writer = svgnetwriter)
```

This is a bit confusing, because the order of arguments changes for `expect_doppleganger` and `writer`, and change meaning as well. The `title` argument to `expect_doppleganger` becomes the `file` argument of `writer` (with `'.svg'` appended). So `writer` needs plot, then filename, then a title (which I skip here). The `file` should *not* have `.svg` on it, as that gets added automatically to `"plot_name"` in `expect_doppleganger`. The `title` argument for `writer` is *only* used for ggplot objects in `vdiffr::print_plot()`.
