---
title: "Atlas of living australia"
author: "Galen Holt"
---

I need to pull some records from the [Atlas of living australia](https://www.ala.org.au/). There's an API. Can I use it so I don't have to do this manually (and can re-do it easily?).

Was going to set up to hit the API directly, but they have an R package (that hits a LOT of these sorts of atlases, which is nice).

```{r}
# library(httr2)
library(galah)
```

Have to register at their website first- *if you're at an Australian university, it's likely you can login with those credentials*- click the AAF button.

```{r}
galah_config(email = "g.holt@deakin.edu.au", atlas = "Australia")

```

## Try it

I'm just following the vignette, but not filtering or selecting

```{r}
lippia <- galah_call() |>
  galah_identify("lippia") |>
  # galah_select(institutionID, group = "basic") |>
  atlas_occurrences()

lippia
```

From the website though, I know that *Phyla nodiflora* comes up as a synonym with way more records. Can I get that sort of info from the API before pulling?

```{r}
all_lippia <- search_taxa('lippia')
all_lippia
```

Looks like no- these come up in the search bar on the website, but not in the above. Oh well, I guess be aware of that.

```{r}
pnode <- search_taxa('phyla nodiflora')
```

## Can I ask for both in one call?

Yeszs

```{r}
lippia_phyla <- galah_call() |>
  galah_identify(c("lippia", 'phyla nodiflora')) |>
  # galah_select(institutionID, group = "basic") |>
  atlas_occurrences()

lippia_phyla
```

And it returns the names, so I can filter.
