---
title: "Bom reference stations"
editor: visual
---

```{r setup}
#| warning: false
#| message: false

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r}
library(ggplot2)
library(sf)
```

Trying to find BOM gauge locations. Found reference stations. It's a simple link, but have to use httr2 to download because there's an error with the `user_agent` if we try to just `download.file`.

Mostly including this here as an example of changing `user_agent`.

[http://www.bom.gov.au/waterdata/](http://www.bom.gov.au/waterdata/z) has a clickable link to what I *want*, but the data is buried in a frame so can't scrape.

The below is because I found a link to reference stations and wanted to see what they were.

### Is there a url for BOM?

It's just a csv, but have to faff about with httr2 and deparsing back to csv because need to pass a user agent or get a 403 error.

```{r}

bom2 <- httr2::request("http://www.bom.gov.au/water/hrs/content/hrs_station_details.csv") |>
  httr2::req_user_agent("md-werp") |> 
  httr2::req_perform() |> 
  httr2::resp_body_string() |> 
  readr::read_csv(skip = 11) |> 
  dplyr::select(site = `Station Name`, 
                gauge = `AWRC Station Number`,  
                owner = `Data Owner Name`, 
                Latitude, Longitude)

```

```{r}
basin <- read_sf(file.path('data','mdb_boundary', 'mdb_boundary.shp'))


```

```{r}
bom2 <- bom2 |>
    # lat an long come in as chr because there is a line for 'undefined'
    dplyr::filter(site != 'undefined') |>
  st_as_sf(coords = c('Longitude', 'Latitude'), crs = 4326) |> 
  st_transform(crs = st_crs(basin))
```

They're not the gauges I'm looking for. Only 457, instead of 6500, and around the edges of the basin.

```{r}

ggplot() + 
  geom_sf(data = basin) +
  geom_sf(data = bom2)
```
