---
title: "Read australia geofabric"
format: html
---

```{r}
library(sf)
library(dplyr)
library(ggplot2)
```

We downloaded and saved the [geofabric](get_geofabric.qmd). Now we want to open and plot some of it.

The folder structure is too complex inside the zip, so have to extract. I'm most interested right now in the river lines.

```{r}
gdbpath <- file.path('data', 'geofabric', 'SH_Network_GDB', 'SH_Network.gdb')
if (!file.exists(gdbpath))
  
unzip(file.path('data', 'geofabric', 'SH_Network_GDB_V3_3.zip'),
      exdir = file.path('data', 'geofabric'))
```

There are a lot of layers, so can't just `st_read`

```{r}
st_layers(gdbpath)
```

The stream lines are big, but just reading doesn't take *too* long.

```{r}
#| cache-lazy: false
rivers <- st_read(gdbpath, layer = 'AHGFNetworkStream')
```

Let's subset to the major rivers to not crash?

```{r}
major_rivers <- rivers |> 
  dplyr::filter(Hierarchy == 'Major')
```

Still `r nrow(major_rivers)`, but we can try to plot it?

```{r}
plot(major_rivers[,'Name'])
```

```{r}
riverlen_plot <- ggplot(major_rivers, aes(color = log(UpstrGeoLn))) +
  geom_sf() +
  theme_void()
```

```{r}
riverlen_plot + theme(legend.position = 'none')
```
