---
title: "Less ugly tables"
echo: true
format:
  html:
    toc: true
    comments:
      hypothesis: true
  docx:
    toc: true
    toc-depth: 2
    prefer-html: true
    # setting these to the html defaults so they don't get jumbled between format
    fig-width: 7
    fig-height: 5
---

These all look fine in html, most things do. But the word doc is where things seem to fall down. Especially if they have lots of text.

```{r}
#| label: tbl-kable
#| tbl-cap: This is a table with kable
head(iris) |> 
  knitr::kable()
```

```{r}
#| eval: false
#| label: tbl-extra
#| tbl-cap: "This is a table with kableExtra. It does not render to word. It used to with `prefer-html: true`, but that workarounds seems not to work anymore."

# head(iris) |> 
#   knitr::kable() |> 
#   kableExtra::kable_styling()
```

```{r}
#| label: tbl-hux
#| tbl-cap: This is a table with huxtable

head(iris) |> 
  huxtable::huxtable() |> 
  huxtable::theme_article()
```

```{r}
#| label: tbl-gt
#| tbl-cap: This is a table with gt

head(iris) |> 
 gt::gt()
```

### Long text

This table is giving me fits elsewhere, let's try it here. It has a common issue of

```{r}
#| label: tbl-raw
#| tbl-cap: This is just printed
#| message: false
texttab <- readr::read_csv('data/component_table.csv', show_col_types = FALSE)

texttab
```

```{r}
#| label: tbl-kable2
#| tbl-cap: This is a table with kable
texttab |> 
  knitr::kable()
```

```{r}
#| label: tbl-hux2
#| tbl-cap: This is a table with huxtable

texttab |> 
  huxtable::huxtable() |> 
  huxtable::theme_article()
```

```{r}
#| label: tbl-gt2
#| tbl-cap: This is a table with gt

texttab |> 
 gt::gt()
```

gt seems to really have a lot of issues here. What if I use divs?

::: {#tbl-gt3}

```{r}
texttab |> 
 gt::gt()
```

This is a table with gt in a div.
:::

## Column widths. 

None of those handle column widths with long text. kable smashes the first col, huxtable and gt smash them all.

Based on some [github](https://github.com/quarto-dev/quarto-cli/issues/7321) [issues](https://github.com/quarto-dev/quarto-cli/issues/7151), does it work if I just don't use a label at all?

Gt no label

```{r}
texttab |> 
 gt::gt()
```

Huxtable no label

```{r}
texttab |> 
  huxtable::huxtable() |> 
  huxtable::theme_article()
```

