---
title: "Less ugly tables"
echo: true
format:
  html:
    toc: true
    comments:
      hypothesis: true
  docx:
    toc: true
    toc-depth: 2
    prefer-html: true
    # setting these to the html defaults so they don't get jumbled between format
    fig-width: 7
    fig-height: 5
    reference-doc: ../default_word_template.docx
---

Tables tend to look fine in html, but when rendering to word they completely fall down. Below, I test a bunch of the standard packages. \*The one that seems to work best is \[flextable::flecxtable()\], which is interesting, since \[huxtable::huxtable()\] claims to build on it but doesn't work well at all.

*NOTE: this doc probably looks fine on the website. The issue is word rendering.*

I'll test two tables- iris, which has narrow columns, and a table with lots of text, which tends to cause the most problems.

```{r}
#| message: false
texttab <- readr::read_csv(file.path('data/component_table.csv'), 
                           show_col_types = FALSE) |> 
  dplyr::select(-tidyselect::last_col())

```

## The solution

Maybe. This exact code works great in another project, and not at all here. I have no idea what's going on.

```{r}
library(flextable)

set_flextable_defaults(font.family = 'Calibri',
                       font.size = 10,
                       table.layout = 'autofit')
```

Use `flextable` with `set_table_properties(layout = 'autofit')` .

```{r}
#| label: tbl-flex
#| tbl-cap: This is a table with flextable
#| message: false

head(iris) |> 
  flextable()

```

```{r}
#| label: tbl-flex-text
#| tbl-cap: This is a table with lots of text in flextable
#| message: false

texttab |> 
  flextable() #|> 
  # set_table_properties(layout = "autofit")

```

Note that there is an issue still if the table prints into a div. I think because divs render in word as tables, and so flextable struggles with what it controls. I'll show that below.

## The others 

These all look fine in html, most things do. But the word doc is where things seem to fall down. Especially if they have lots of text.

```{r}
#| label: tbl-kable
#| tbl-cap: This is a table with kable
head(iris) |> 
  knitr::kable()
```

```{r}
#| eval: false
#| label: tbl-extra
#| tbl-cap: "This is a table with kableExtra. It does not render to word. It used to with `prefer-html: true`, but that workarounds seems not to work anymore."

# head(iris) |> 
#   knitr::kable() |> 
#   kableExtra::kable_styling()
```

```{r}
#| label: tbl-hux
#| tbl-cap: This is a table with huxtable

head(iris) |> 
  huxtable::huxtable() |> 
  huxtable::theme_article()
```

```{r}
#| label: tbl-gt
#| tbl-cap: This is a table with gt

head(iris) |> 
 gt::gt()
```

## Long text

This table is giving me fits elsewhere, let's try it here. It has a common issue of wrapping words.

```{r}
#| label: tbl-raw
#| tbl-cap: This is just printed

texttab
```

```{r}
#| label: tbl-kable2
#| tbl-cap: This is a table with kable
texttab |> 
  knitr::kable()
```

```{r}
#| label: tbl-hux2
#| tbl-cap: This is a table with huxtable

texttab |> 
  huxtable::huxtable() |> 
  huxtable::theme_article()
```

```{r}
#| label: tbl-gt2
#| tbl-cap: This is a table with gt

texttab |> 
 gt::gt()
```

gt seems to really have a lot of issues here. What if I use divs?

::: {#tbl-gt3}
```{r}
texttab |> 
 gt::gt()
```

This is a table with gt in a div.
:::

## Column widths.

None of those handle column widths with long text. kable smashes the first col, huxtable and gt smash them all.

Based on some [github](https://github.com/quarto-dev/quarto-cli/issues/7321) [issues](https://github.com/quarto-dev/quarto-cli/issues/7151), does it work if I just don't use a label at all?

Gt no label

```{r}
texttab |> 
 gt::gt()
```

Huxtable no label

```{r}
texttab |> 
  huxtable::huxtable() |> 
  huxtable::theme_article()
```

## Divs

Flextable struggles in divs. The best we can do is to set `width = 1` to at least make all the cols the same width and use the whole space. Though usually what i end up doing is rendering those tables outside a div and pasting in.

::: {#divtable style="color: gray"}
```{r}
#| label: tbl-flex-div
#| tbl-cap: This is a flextable in a div with `width = 1`

texttab |> 
  flextable::flextable() |> 
  flextable::set_table_properties(layout = "autofit", width = 1)


```

If we don't set widths, it's smooshed, even with autofit.

```{r}
#| label: tbl-flex-div-nowidth
#| tbl-cap: This is a flextable in a div with `width = 1`

texttab |> 
  flextable::flextable() |> 
  flextable::set_table_properties(layout = "autofit")
```
:::
