---
title: "R-py type passing"
author: "Galen Holt"
---

```{r setup}
#| warning: false
#| message: false

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

## Passing types

I have a python function that takes dicts and lists as arguments. How do I pass those from R? We can't just assign them to a variable in R, because those formats don't work- e.g. we cannot create the lists and dicts in R to pass.

```{r}
#| eval: false
outputType = ['summary', 'all']

allowance ={'minThreshold': MINT, 'maxThreshold': MAXT, 'duration': DUR, 'drawdown': DRAW}
```

So, let's write a function to tell me the type of what I'm passing and try a few things.

```{python}
def test_type(testarg):
  
  return(type(testarg))
  
```

Is a named list a dict or a list? what about just a `c()`? Is that a list?

```{r}
rlist <- list(dict1 = 100, dict2 = 'testing')
rc <- c(100, 50)
```

The named list is a dict

```{r}
py$test_type(rlist)
```

The c() is a list- but see below- this fails if it's length-one

```{r}
py$test_type(rc)
```

An unnamed list is a list

```{r}
rlistu = list(100, 'testunname')
py$test_type(rlistu)
```

That is useful, since creating a length-one list doesn't work with single values or `c()` wrapping single values

```{r}
rone = 100
py$test_type(rone)
```

```{r}
ronec = c('testingc')
py$test_type(ronec)
```

We *do* get a length-one list with an unnamed list of length 1.

```{r}
ronel = list(100)
py$test_type(ronel)
```
